### 2010-07-23 papers/icfp-2010-desc/paper_desc_levitation.tex
Desc@ICFP: a slightly less grungy Skyhook.

In this section, we have seen how to \emph{levitate}
descriptions. From this unusual situation, several questions naturally
arise. First, we have blurred the line between the implementation of
data-types and their reflection in themselves. Second, we have abused
the paradoxical nature of $\Bhab{\Set}{\Set}$ to flatten the hierarchy
of descriptions. We shall now be more precise about these points.

@@

Let us first clarify the status of the implementation. The kit for
making datatypes is summed up in Table~\ref{tab:sumup-operators}. For
each operation, we describe its role and its status, making clear
which components are self-described and which ones are actually
implemented.

@@

To tackle the apparent circularity of our construction, let us
consider how it bottoms out in a stratified setting. Using the
informal level annotations of Section~\ref{sec:universe-desc}, we have
that \(\Bhab{\DescD_n}{\Desc_{n+1}}\), hence \(\Bhab{\Desc_n \mapsto
  \Mu{\DescD_n}}{\Set_{n+1}}\). Therefore, there is no circularity in
the description of descriptions: assuming an hardwired $\Desc_N$
universe, we can build the tower of descriptions down to
$\Desc_0$\footnote{We have modelled this process in Agda, hardwiring
  $\Desc_{42}$ in $\Set_{43}$, then going down. This model does not
  rely on dependent pattern-matching, induction-recursion, nor set
  polymorphism.}. Hence, in a stratified system, the `self-encoded'
nature of $\Desc$ appears only in a set polymorphic sense: the
principal type of the levitating description generalizes to the type
of $\Desc$ itself\footnote{This is witnessed in our set polymorphic
  model in Agda. Set polymorphism is an experimental feature of
  Agda. Although not backed by theoretical works, it is a conservative
  extension of Agda that is generally accepted as free of paradox.}.

>>>>

In this section, we have seen how to \emph{levitate}
descriptions. From this unusual situation, several questions naturally
arise. First, we have blurred the line between the implementation of
data-types and their self-description. Second, we have abused the
paradoxical nature of $\Bhab{\Set}{\Set}$ to flatten the hierarchy of
descriptions. We shall now be more precise about these points.

@@

Let us first clarify the status of the implementation. The kit for
making datatypes is presented in Table~\ref{tab:sumup-operators}. For
each operation, we describe its role and its status, making clear
which components are self-described and which ones are actually
implemented.

@@

To demystify the apparent circularity of our construction, let us
consider how it bottoms out in a stratified setting. Using the
informal level annotations of Section~\ref{sec:universe-desc}, we have
that \(\Bhab{\DescD^n}{\Desc^{n+1}}\), hence \(\Bhab{\Desc^n \mapsto
  \Mu{\DescD^n}}{\Set^{n+1}}\). Assuming an hardwired $\Desc^N$
universe, we can build the tower of descriptions down to $\Desc^0$:
description of descriptions forms a `spiral', rather than a
circle \footnote{We have modelled this process in Agda, hardwiring
  $\Desc^{42}$ in $\Set^{43}$, then going down. This model does not
  rely on dependent pattern-matching, induction-recursion, nor set
  polymorphism.}. Hence, in a stratified system, the `self-encoded'
nature of $\Desc$ appears only in a set polymorphic sense: the
principal type of the levitating description generalizes to the type
of $\Desc$ itself\footnote{This is witnessed in our set polymorphic
  model in Agda. Set polymorphism is an experimental feature of
  Agda. Although not backed by theoretical works, it is a conservative
  extension of Agda that is generally accepted as free of paradox.}.
